---
title: "Ridesharing Project"
output: html_document
editor_options: 
  chunk_output_type: console
---

##### *Nate Wagner*
##### *Olivia Roberts*
##### *Dominic Ventura*

<br>

## Introduction

This project conducts statistical analysis on Uber and Lyft price estimates, and seeks to find associations between price and other variables. The data comes from [Kaggle](https://www.kaggle.com/ravi72munde/uber-lyft-cab-prices), and was collected via Uber and Lyft api queries and corresponding weather conditions during a few weeks in November-December in 2018 by this [user](https://github.com/ravi72munde). Inspiration for cleaning and combining the rides and weather table came from [here](https://www.kaggle.com/tangyx233/uber-lyft-cab-prices).


```{r message=FALSE, warning=FALSE, include=FALSE}
library(readr)
library(tidyverse)
library(knitr)
library(reshape)
library(lubridate)
library(gridExtra)
library(sqldf)
#library(sjPlot)
#library(sjmisc)
#library(sjlabelled)
library(pscl)
library(kableExtra)
library(jtools)
library(stargazer)
library(formattable)
#library(data.table)
#install.packages("stargazer")
#install.packages("jtools")
#install.packages("pscl")
#install.packages("sjPlot")
#install.packages("sjmisc")
#install.packages("sjlabelled")
```


```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
weather <- read_csv("uber-lyft-cab-prices/weather.csv")
rides <- read_csv("uber-lyft-cab-prices/cab_rides.csv")
```


```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# To replace missing values of rain with 0
weather$rain[is.na(weather$rain)] <- 0
```


```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# To fix the timestamp:
rides$timestamp <- as.POSIXct(rides$time_stamp/1000,origin='1970-01-01',tz='America/New_York')
rides$Date <- as.Date(rides$timestamp)
rides$Time <- format(rides$timestamp,"%H:%M:%S")
rides$weekdays <- weekdays(rides$Date)

weather$timestamp<-as.POSIXct(weather$time_stamp,origin='1970-01-01',tz='America/New_York')
weather$Date<-as.Date(weather$timestamp)
weather$Time <- format(weather$timestamp,"%H:%M:%S")
```


```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# Merge datasets
weather$hour<-hour(hms(weather$Time))
weather$merge<-paste(weather$location,weather$Date,weather$hour)
rides$hour<-hour(hms(rides$Time))
rides$merge<-paste(rides$destination,rides$Date,rides$hour)

weatherDF <- sqldf("select AVG(temp) as temp_avg,AVG(clouds) as clouds_avg,AVG(pressure) as pressure_avg,AVG(rain) as rain_avg,AVG(humidity) as humidity_avg,AVG(wind) as wind_avg,merge from weather group by merge")

mergeDF<-merge(rides,weatherDF,by='merge')
#drop redundant columns

```


```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# Add time of day column and remove 'taxi'
mergeDF$TOD <- with(mergeDF,  ifelse(hour >= 5 & hour<=11, "morning",
                            ifelse(hour>11 & hour<=16, "evening", "night")))


mergeDF <- mergeDF %>% filter(name != 'Taxi')
mergeDF <- mergeDF %>% filter(name != 'WAV')



levels(as.factor(mergeDF$name[mergeDF$cab_type == "Uber"]))
levels(as.factor(mergeDF$name[mergeDF$cab_type == "Lyft"]))

#mergeDF <- mergeDF[,-25]


# to clean up name column:
n <- nrow(mergeDF)
for (x in 1:n) {
  if (mergeDF$name[x] == 'Lyft' | mergeDF$name[x] == 'UberX') {
    mergeDF$type[x] <- 'Basic'
  } 
  if (mergeDF$name[x] == 'Lux Black XL' | mergeDF$name[x] == 'Black SUV') {
    mergeDF$type[x] <- 'LuxXL'
  }
  if (mergeDF$name[x] == 'Lyft XL' | mergeDF$name[x] == 'UberXL') {
    mergeDF$type[x] <- 'XL'
  }
  if (mergeDF$name[x] == 'Shared' | mergeDF$name[x] == 'UberPool') {
    mergeDF$type[x] <- 'Shared'
  }
  if (mergeDF$name[x] == 'Black' | mergeDF$name[x] == 'Lux Black' | mergeDF$name[x] == 'Lux') {
    mergeDF$type[x] <- 'Lux'
  }

}


write.csv(mergeDF, file = "CleanedRideSharing.csv")

```


```{r message=FALSE, warning=FALSE, include=FALSE}
mergeDF <- read_csv("CleanedRideSharing.csv")
```


```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
## Check data for completness:
ridesNA <- mergeDF %>%
  select(everything()) %>%
  summarise_all(funs(sum(is.na(.))))

PropRidesNA <- (ridesNA/nrow(rides))*100
kable(PropRidesNA, format = "markdown")
```

<br>

## Exploratory Data Analysis

Our variable of interest is price, which ranges from $2.50 to $97.50.

<br>

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary1 <- as.data.frame(as.array(summary(mergeDF$price)))
names(summary1) <- c("-   Price", "-   Value")
kable(summary1, type = "Markdown")
#stargazer(summary1, type = "html")
```

<br>

The distribution of price is about the same for both Uber and Lyft
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(mergeDF, aes(price, color = cab_type)) +
  geom_histogram(fill="white", alpha=0.5, position="identity", alpha=0.5) +
  theme_bw()
```

```{r eval=FALSE, include=FALSE}
# boxplot of price by cab type
ggplot(mergeDF, aes(as.factor(cab_type), price)) +
  geom_boxplot(varwidth=T, fill="plum") +
  labs(x = "Cab Type") +
  theme_bw()
```


#### How does price vary by cab type?
```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
levels(as.factor(mergeDF$name[mergeDF$cab_type == "Uber"]))
levels(as.factor(mergeDF$name[mergeDF$cab_type == "Lyft"]))
```
<br>
Here is the min, max, median and mean price of a trip broken down by the vehicle type

<br>
```{r echo=FALSE, message=FALSE, warning=FALSE}
price1 <- mergeDF %>%
            filter(cab_type == "Uber") %>%
            group_by(type) %>% 
            summarise(min(price), max(price), median(price), mean(price))
names(price1) <- c("-   Uber Ride Type", "-   Uber Min Price", "-   Uber Max Price", "-   Uber Median Price", "-   Uber Mean Price")

price2 <- mergeDF %>%
            filter(cab_type == "Lyft") %>%
            group_by(type) %>% 
            summarise(min(price), max(price), median(price), mean(price))
names(price2) <- c("-   Lyft Ride Type", "-   Lyft Min Price", "-   Lyft Max Price", "-   Lyft Median Price", "-   Lyft Mean Price")

kable(price1, type = "markdown")
kable(price2, type = "markdown")
```
<br>

There are 5 different vehicle types for Uber (Black, Black SUV, Uber Pool, UberX, and UberXL) and there are 6 different vehicle types for Lyft (Lux, Lux Black, Lux Black XL, Lyft, Lyft XL and Shared) So to make proper comparisons, we made the following comparisons:

<br>

* Lyft is comparable to UberX
* Lux Black XL is comparable to Black SUV
* Lyft XL is comparable to UberXL
* Shared is comparable to Uber Pool
* Black is comparable to Lux Black

<br>

We also combined Lux with Lux Black because they are essentially the same vehicle type minus some small difference. 

<br>

Here we can clearly see how the distribution of price changes depending on the ride type. 

<br>

```{r echo=FALSE, message=FALSE, warning=FALSE}
Uber.price <- ggplot(mergeDF %>% filter(cab_type == "Uber") %>% group_by(type), aes(reorder(as.factor(type), price, FUN = median), price, fill = type)) +
  geom_boxplot(position=position_dodge(1)) +
  labs(title = "Uber", x = " ") +
  theme_bw()

Lyft.price <- ggplot(mergeDF %>% filter(cab_type == "Lyft") %>% group_by(type), aes(reorder(as.factor(type), price, FUN = median), price, fill = type)) +
  geom_boxplot(position=position_dodge(1)) +
  labs(title = "Lyft", x = " ") +
  theme_bw()

grid.arrange(Uber.price, Lyft.price, nrow = 2)

```


#### Distance vs Price
<br>
The distribution of price is roughly bimodal and a little skewed right
<br>
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(mergeDF, aes(distance)) +
  geom_histogram(col = "black", fill="lightblue", alpha=0.5, position="identity", alpha=0.5) +
  theme_bw()
```

As we can see this by scatter plot, Lyft has a lower base price and also a higher price ceiling.
<br>
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(mergeDF, aes(distance, price)) +
  geom_point(alpha = 0.05, aes(col = cab_type)) +
  labs(title = "Uber") +
  theme_bw()
```



```{r echo=FALSE, message=FALSE, warning=FALSE}
#So far the comparisons have been on the base price between Uber and Lyft. We have data on surge prices for Lyft, but not Uber.
mergeDF <- mutate(mergeDF, SurgePrice = surge_multiplier * price)
```



Here is a scatter plot of price vs distance broken down by ride type for both Uber and Lyft. When broken down by ride type, it becomes clear there is potentially some strong positive linear relationship between price and distance. 
<br>
```{r echo=FALSE, message=FALSE, warning=FALSE}
DvP1 <- ggplot(mergeDF %>% filter(cab_type == "Uber" & name != "Taxi"), aes(distance, price)) +
  geom_point(alpha = 0.05, aes(col = name)) +
  labs(title = "Uber") +
  theme_bw()

DvP2 <- ggplot(mergeDF %>% filter(cab_type == "Lyft" & name != "Taxi"), aes(distance, price)) +
  geom_point(alpha = 0.05, aes(col = name)) +
  labs(title = "Lyft") +
  theme_bw() 

DvP1 
DvP2
```

```{r message=FALSE, warning=FALSE, include=FALSE}
cor(mergeDF %>%
      filter(cab_type == "Lyft" & name == "Lyft") %>%
      select(distance),
    mergeDF %>%
      filter(cab_type == "Lyft" & name == "Lyft") %>%
      select(price))
cor(mergeDF %>%
      filter(cab_type == "Lyft" & name == "Lux Black") %>%
      select(distance),
    mergeDF %>%
      filter(cab_type == "Lyft" & name == "Lux Black") %>%
      select(price))
cor(mergeDF %>%
      filter(cab_type == "Lyft" & name == "Lux Black XL") %>%
      select(distance),
    mergeDF %>%
      filter(cab_type == "Lyft" & name == "Lux Black XL") %>%
      select(price))
cor(mergeDF %>%
      filter(cab_type == "Lyft" & name == "Lux") %>%
      select(distance),
    mergeDF %>%
      filter(cab_type == "Lyft" & name == "Lux") %>%
      select(price))
cor(mergeDF %>%
      filter(cab_type == "Lyft" & name == "Lyft XL") %>%
      select(distance),
    mergeDF %>%
      filter(cab_type == "Lyft" & name == "Lyft XL") %>%
      select(price))
cor(mergeDF %>%
      filter(cab_type == "Lyft" & name == "Shared") %>%
      select(distance),
    mergeDF %>%
      filter(cab_type == "Lyft" & name == "Shared") %>%
      select(price))
```
<br>
```{r echo=FALSE, message=FALSE, warning=FALSE}
cor_lyft <- matrix(c(0.7798634,"","","","","","",0.8132575,"","","","","","",0.6973414,"","","","","","",0.8366301,"","","","","","",0.8144453,"","","","","","",0.6420404), ncol = 6, byrow=TRUE)
colnames(cor_lyft) <- c("   Lyft", "Lux Black", " Lux Black XL", "   Lux", "Lyft XL", "Shared")
rownames(cor_lyft) <- c("Lyft", "Lux Black", "Lux Black XL", "Lux", "Lyft XL", "Shared")
cor_lyft <- data.frame(
  "Price Vs Distance" = c("Lyft", "Lux Black", "Lux Black XL", "Lux", "Lyft XL", "Shared"),
  Lyft = c(0.7798634,"","","","",""),
  LuxBlack = c("",0.8132575,"","","",""),
  LuxBlackXL = c("","",0.6973414,"","",""),
  Lux = c("","","",0.8366301,"",""),
  LyftXL = c("","","","", 0.8144453,""),
  Shared = c("","","","","",0.6420404))
formattable(cor_lyft,
            align = c("l",rep("r", NCOL(cor_lyft) - 1)),
            list(`"Price Vs Distance"` = formatter("span", style = ~ style(color = "grey", font.weight = "bold")),
                 `Lyft` = color_bar("lightgreen"),
                 `LuxBlack` = color_bar("lightgreen"),
                 `LuxBlackXL` = color_bar("lightgreen"),
                 `Lux` = color_bar("lightgreen"),
                 `LyftXL` = color_bar("lightgreen"),
                 `Shared` = color_bar("lightgreen")))
```


```{r message=FALSE, warning=FALSE, include=FALSE}
cor(mergeDF %>%
      filter(cab_type == "Uber" & name == "Black") %>%
      select(distance),
    mergeDF %>%
      filter(cab_type == "Uber" & name == "Black") %>%
      select(price))
cor(mergeDF %>%
      filter(cab_type == "Uber" & name == "Black SUV") %>%
      select(distance),
    mergeDF %>%
      filter(cab_type == "Uber" & name == "Black SUV") %>%
      select(price))
cor(mergeDF %>%
      filter(cab_type == "Uber" & name == "UberPool") %>%
      select(distance),
    mergeDF %>%
      filter(cab_type == "Uber" & name == "UberPool") %>%
      select(price))
cor(mergeDF %>%
      filter(cab_type == "Uber" & name == "UberX") %>%
      select(distance),
    mergeDF %>%
      filter(cab_type == "Uber" & name == "UberX") %>%
      select(price))
cor(mergeDF %>%
      filter(cab_type == "Uber" & name == "UberXL") %>%
      select(distance),
    mergeDF %>%
      filter(cab_type == "Uber" & name == "UberXL") %>%
      select(price))
cor(mergeDF %>%
      filter(cab_type == "Uber" & name == "WAV") %>%
      select(distance),
    mergeDF %>%
      filter(cab_type == "Uber" & name == "WAV") %>%
      select(price))
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
cor_uber <- matrix(c(0.9131188,"","","","","","",0.8582338,"","","","","","",0.670947,"","","","","","",0.7350737,"","","","","","",0.7859246,"","","","","","",0.7353036), ncol = 6, byrow=TRUE)
colnames(cor_uber) <- c("   Black", "Black SUV", " UberPool", "   UberX", "UberXL", "WAV")
rownames(cor_uber) <- c("Black", "Black SUV", "UberPool", "UberX", "UberXL", "WAV")
# cor_uber <- as.table(cor_uber)
#names(attributes(cor_uber)$dimnames) <- c("  Price","                          Distance")
cor_uber <- data.frame(
  "Price Vs Distance" = c("Black", "Black SUV", "UberPool", "UberX", "UberXL", "WAV"),
  Black = c(0.7798634,"","","","",""),
  BlackSUV = c("",0.8132575,"","","",""),
  UberPool = c("","",0.6973414,"","",""),
  UberX = c("","","",0.8366301,"",""),
  UberXL = c("","","","", 0.8144453,""),
  WAV = c("","","","","",0.6420404))
formattable(cor_uber,
            align = c("l",rep("r", NCOL(cor_lyft) - 1)),
            list(`"Price Vs Distance"` = formatter("span", style = ~ style(color = "grey", font.weight = "bold")),
                 `Black` = color_bar("lightgreen"),
                 `BlackSUV` = color_bar("lightgreen"),
                 `UberPool` = color_bar("lightgreen"),
                 `UberX` = color_bar("lightgreen"),
                 `UberXL` = color_bar("lightgreen"),
                 `WAV` = color_bar("lightgreen")))
```
<br>
To confirm, there are high correlations between distance and price when broken down by ride type here.
<br>
<br>

#### Price on weekdays for Lyft

So far the comparisons have been on the base price between Uber and Lyft. We only have data on surge prices for Lyft, but not Uber. So here we make comparisons based off Lyft prices (including surge) for Lyft only.

There doesn't seem to be an effect of the day of the week on median Surge Prices according to the data, and the weeks that it was collected. This is obviously surprising, we would expect prices to be higher on Fridays or even weekends. 

<br>
```{r echo=FALSE, message=FALSE, warning=FALSE}
week.lyft <- ggplot(mergeDF %>% filter(cab_type == "Lyft") %>% group_by(weekdays), aes(weekdays, SurgePrice, fill = weekdays))+
  geom_boxplot(position=position_dodge(1)) +
  labs(title = "Lyft", x = " ") +
  theme_bw()

week.lyft
```


```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# Price on destination:
dest.uber <- ggplot(mergeDF %>% filter(cab_type == "Uber") %>% group_by(destination), aes(destination, price, fill = destination))+
  geom_boxplot(position=position_dodge(1)) +
  labs(title = "Uber", x = " ") +
  theme_bw()

dest.lyft <- ggplot(mergeDF %>% filter(cab_type == "Lyft") %>% group_by(destination), aes(destination, price, fill = destination))+
  geom_boxplot(position=position_dodge(1)) +
  labs(title = "Lyft", x = " ") +
  theme_bw()

grid.arrange(dest.uber, dest.lyft)
```
<br>

#### Price on time of day

There also doesn't seem to be much difference in the median of surge price when it is broken down by the time of day, according to our data and the weeks it was collected. 
<br>


```{r echo=FALSE, message=FALSE, warning=FALSE}
tod.lyft <- ggplot(mergeDF %>% filter(cab_type == "Lyft") %>% group_by(TOD), aes(TOD, SurgePrice, fill = TOD)) + geom_boxplot(position=position_dodge(1)) + labs(title = "Lyft", x = " ") + theme_bw()

tod.lyft
```



#### Mean base price of an Uber trip 
<br>
```{r message=FALSE, warning=FALSE, include=FALSE}
price <- mergeDF %>%
  filter(cab_type == "Uber") %>%
  select(price)
mean.uber <- mergeDF %>%
  filter(cab_type == "Uber") %>%
  select(price) %>%
  summarize(average = mean(price))
sd.uber <- mergeDF %>%
  filter(cab_type == "Uber") %>%
  select(price) %>%
  summarize(average = sd(price))
n <- mergeDF %>%
  filter(cab_type == "Uber") %>%
  summarize(sample_size = n())
price <- as.vector(price)
#is.vector(price)
mean.uber <- as.integer(mean.uber)
sd.uber <- as.integer(sd.uber)
n<- as.integer(n)
CI<- c(mean.uber - (1.95*(sd.uber/sqrt(n))) , mean.uber + (1.95*(sd.uber/sqrt(n))))
#print(CI)
```
The estimated mean base price of an Uber trip in Boston:
$$ \begin{equation} 
\bar{x} = 16 
\end{equation} $$

95% Confidence Interval:
$$ \begin{equation} 
(15.97, 16.03) 
\end{equation} $$

We are 95% confident that the true population mean base price of an Uber Trip in Boston is between $15.97 and $16.03

Here we conduct a hypothesis test to see whether the mean base price of an Uber trip in Boston differs from $10.
The null and alternative hypothesis are as follows:

$$ H_{0}: \mu = 10 $$





$$ H_{A}: \mu \neq 10 $$

```{r echo=FALSE, message=FALSE, warning=FALSE}
Ttest <- t.test(price$price, mu = 10, conf.level = 0.95)
Ttest
eff.size<- (mean.uber - 10)/sd.uber
#print(eff.size)
```
The conclusion is to reject the null hypothesis that the true population mean of base price for Uber is equal to $10 at all alpha levels. The p-value is essentially zero due to the large sample size and the effect size of 0.75 confirms practical significance. The confidence interval also agrees with our conclusion in that it does not contain $10.


<br>
<br>

#### Proportion of Uber X trips
```{r echo=FALSE, message=FALSE, warning=FALSE}
x <- mergeDF %>% filter(name == "UberX") %>% count()
n <- nrow(mergeDF %>% filter(cab_type == "Uber"))
p.hat <- x/n
#prop.test(as.integer(x), n, 0.30)
se <- sqrt(p.hat*(1-p.hat)/n)
CI <- c(p.hat - (1.95*se) , p.hat + (1.95*se))
```
The estimated proportion of Uber X trips in Boston is
$$ \begin{equation} 
\hat{p} = 0.39
\end{equation} $$



95% Confidence Interval:
$$ \begin{equation} 
(0.384, 0.386) 
\end{equation} $$


We are 95% confident that the true population proportion of Uber X trips in Boston is between 0.384 and 0.386


Here we conduct a hypothesis test to see whether the proportion of Uber X trips in Boston is different from 0.30. 
The null and alternative hypothesis are as follows:

$$ H_{0}: p = 0.30 $$



$$ H_{0}: p \neq 0.30$$


```{r echo=FALSE, message=FALSE, warning=FALSE}
prop.test(as.integer(x), n, 0.30)
effect_size <- 2* abs(asin(sqrt(x/n))-asin(sqrt(0.30)))
```
The conclusion is to reject the null hypothesis that the true population proportion of Uber X trips is equal to 0.30 at all alpha levels. The p-value is essentially zero due to the large sample size and the effect size of 0.18 says there is little practical significance. The confidence interval also agrees with our conclusion in that it does not contain 0.30.



<br>
<br>


#### Is the mean estimated price different between Uber and Lyft?


The null and alternative hypothesis are as follows

$$ H_{0}: \mu_{U}-\mu_{L} = 0 $$

$$ H_{A}: \mu_{U}-\mu_{L} \neq 0 $$
```{r echo=FALSE, message=FALSE, warning=FALSE}
t.test(mergeDF$price[mergeDF$cab_type == "Uber"], mergeDF$price[mergeDF$cab_type == "Lyft"])

x1 <- mean(mergeDF$price[mergeDF$cab_type == "Uber"])
x2 <- mean(mergeDF$price[mergeDF$cab_type == "Lyft"])
s1 <- sd(mergeDF$price[mergeDF$cab_type == "Uber"])
s2 <- sd(mergeDF$price[mergeDF$cab_type == "Lyft"])
n1 <- length(mergeDF$price[mergeDF$cab_type == "Uber"])
n2 <- length(mergeDF$price[mergeDF$cab_type == "Lyft"])


effect_size <- abs((x1 - x2)) / sqrt((s1 + s2)/2)
```
The conclusion is to reject the null hypothesis that the true population difference in mean estimated price between Uber and Lyft is equal to 0 at all alpha levels. The p-value is essentially zero due to the large sample size and the effect size of 0.11 says there may be little practical significance. The confidence interval also agrees with our conclusion in that it does not contain 0. Additionally, the mean difference shows that Lyft has higher prices on average. 




```{r message=FALSE, warning=FALSE, include=FALSE}
# remove outliers of price and distance
lower <- quantile(mergeDF$price, .25) - 1.5*IQR(mergeDF$price)
upper <- quantile(mergeDF$price, .75) + 1.5*IQR(mergeDF$price)

# there are 5566 rows with price greater than 1.5xIQR + 75% quantile
mergeDF %>% filter(price > upper) %>% count()

# remove those values
rmOutliers <- mergeDF %>% filter(price < upper)

upper.d <- quantile(rmOutliers$distance, .75) + 1.5*IQR(rmOutliers$distance)

# there are 617 rows with distance greater than 1.5xIQR + 75% quantile
rmOutliers %>% filter(distance > upper.d) %>% count()

# removed those values
rmOutliers <- rmOutliers %>% filter(distance < 2.90 + 1.5*(IQR(rmOutliers$distance)))


levels(as.factor(mergeDF$type))
```

<br>
<br>
<br>

The first model just regresses price on cab type. 
<br>
*Model 1*:
$$ \begin{equation}
\hat{Price}_i = \hat{\beta}_0 + \hat{\beta}_1 Uber_i
\end{equation} $$
```{r message=FALSE, warning=FALSE, include=FALSE, results='asis'}
# base model
model1 <- lm(price ~ cab_type, data = mergeDF)
#summary(model1)
#stargazer(model1, type = "html")
```


The second model controls for distance. 
<br>
*Model 2*:
$$ \begin{equation}
\hat{Price}_i = \hat{\beta}_0 + \hat{\beta}_1 Uber_i +  \hat{\beta}_1 Dist_i
\end{equation} $$
```{r message=FALSE, warning=FALSE, include=FALSE, results='asis'}
model2 <- lm(price ~ cab_type + distance, data = mergeDF)
```


The Third model controls for distance and the ride type. 
<br>
*Model 3*:
$$ \begin{align*}
\hat{Price}_i = \hat{\beta}_0 + \hat{\beta}_1 Uber_i + \hat{\beta}_2 Dist_i  +\sum_{j=1}^{i}\hat{\beta_3}_j {Vehicle_i}_j  \\ 
\end{align*} $$


```{r message=FALSE, warning=FALSE, include=FALSE, results='asis'}
model3 <- lm(price ~ cab_type + distance + as.factor(type), data = mergeDF)
summary(model3)
```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#model4 <- lm(price ~ cab_type + distance + as.factor(type) + temp_avg + , data = mergeDF)
#summary(model4)
```



```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

stargazer::stargazer(model1, model2, model3, type = "html", align = TRUE, font.size = "small", digits = 2, column.labels = c("-    Model 1    ", "-   Model 2    ", "-   Model 3    "), title = "Regression Results", dep.var.labels=c("Price", "Price", "Price"), covariate.labels=c("Uber", "Distance", "Lux", "Lux XL", "Shared", "XL"),  omit.stat=c("LL","ser","f"),report = "vcp*", column.sep.width = "40")
#testM <- lm(price ~ name, data = rmOutliers)
#summary(testM)
#plot(model3,which = 1)
#qqnorm(model3$residuals)
#quantile(model3$residuals)
#which.max(model3$residuals)
#hist(model3$residuals)
#quantile(model2$residuals)
#plot(model2$residuals)
#qqnorm(model2$residuals)

```

<br>

To also confirm that Time of day as well as the day of the week has no effect on surge prices we ran the following regression:

<br>

*Surge Price Model*:
$$ \begin{align*}
\hat{Price}_i = \hat{\beta}_0 + \hat{\beta}_1 Uber_i + \hat{\beta}_2 Dist_i  +\sum_{j=1}^{i}\hat{\beta_3}_j {Vehicle_i}_j +\sum_{j=1}^{i}\hat{\beta_4}_j {TOD_i}_j +\sum_{j=1}^{i}\hat{\beta_5}_j {Weekday_i}_j\\  
\end{align*} $$


```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
SurgeModel <- lm(SurgePrice ~ distance + type + TOD + weekdays, data = mergeDF %>% filter(cab_type == "Lyft"))

stargazer(SurgeModel, type = "html", digits = 2, title = "Regression Results", dep.var.labels = "Price", covariate.labels=c("Distance", "Lux", "Lux XL", "Shared", "XL", "Morning", "Night", "Monday", "Saturday", "Sunday", "Thursday", "Tuesday", "Wednesday"), omit.stat=c("LL","ser","f"), report = "vcp*")

```

<br>
<br>

<br>

#### Conclusion

The effect of Uber on the first model is -0.35, which is just the difference in means of estimated price between Uber and Lyft. This suggest that the base price of Uber is on average -$0.35 less than Lyft's base price. Adding distance to the second model doesn't do much to the coeffcient on Uber, but distance is statistically significant. Now when we control for the vehicle type of the cab, that's when we see the difference in means switch signs, suggesting that controling for distance and vehicle type, Uber is on average $0.24 more than Lyft's base price. When looking at the surge price model, we do confirm there is no effect of the time of day and the day of the week on surge price estimates in Boston, which is consistent with our boxplots above. 

<br>
<br>
<br>
