---
title: "Statistics Project"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(readr)
library(tidyverse)
library(knitr)
library(reshape)
library(lubridate)
library(gridExtra)
```


```{r}
weather <- read_csv("uber-lyft-cab-prices/weather.csv")
rides <- read_csv("uber-lyft-cab-prices/cab_rides.csv")
```


# Check data for completness:
```{r}
ridesNA <- rides %>%
  select(everything()) %>%
  summarise_all(funs(sum(is.na(.))))

PropRidesNA <- (ridesNA/nrow(rides))*100
kable(PropRidesNA, format = "markdown")

weatherNA <- weather %>%
  select(everything()) %>%
  summarise_all(funs(sum(is.na(.))))

PropWeatherNA <- (weatherNA/nrow(weather))*100
kable(PropWeatherNA, format = "markdown")
```
Roughly 8% of values from price column missing in rides, the rest are complete. 85% of values from rain missing in weather table, the rest are complete. It's proabably that the NA values in the rain column just correspond to 0% chance of rain. 

```{r}
# To replace missing values of rian with 0
weather$rain[is.na(weather$rain)] <- 0
```

# To fix the timestamp:
```{r}
rides$timestamp <- as.POSIXct(rides$time_stamp/1000,origin='1970-01-01',tz='America/New_York')
rides$Date <- as.Date(rides$timestamp)
rides$Time <- format(rides$timestamp,"%H:%M:%S")
rides$weekdays <- weekdays(rides$Date)

weather$timestamp<-as.POSIXct(weather$time_stamp,origin='1970-01-01',tz='America/New_York')
weather$Date<-as.Date(weather$timestamp)
weather$Time <- format(weather$timestamp,"%H:%M:%S")
```

# Merge datasets
```{r}
weather$hour<-hour(hms(weather$Time))
weather$merge<-paste(weather$location,weather$Date,weather$hou)
rides$hour<-hour(hms(rides$Time))
rides$merge<-paste(rides$destination,rides$Date,rides$hour)
```


# Variable of interest
```{r}
summary(rides$price)

ggplot(rides, aes(price, color = cab_type)) +
  geom_histogram(fill="white", alpha=0.5, position="identity", alpha=0.5) +
  theme_bw()

ggplot(rides, aes(as.factor(cab_type), price)) +
  geom_boxplot(varwidth=T, fill="plum") +
  labs(x = "Cab Type") +
  theme_bw()
```


# How does price vary by cab type?
```{r}
levels(as.factor(rides$name[rides$cab_type == "Uber"]))
levels(as.factor(rides$name[rides$cab_type == "Lyft"]))


rides %>%
  filter(cab_type == "Uber") %>%
  group_by(name) %>% 
  summarise(min(price), max(price), median(price), mean(price))

rides %>%
  filter(cab_type == "Lyft") %>%
  group_by(name) %>% 
  summarise(min(price), max(price), median(price), mean(price))


Uber.price <- ggplot(rides %>% filter(cab_type == "Uber" & name != "Taxi") %>% group_by(name), aes(reorder(as.factor(name), price, FUN = median), price, fill = name)) +
  geom_boxplot(position=position_dodge(1)) +
  labs(title = "Uber", x = " ") +
  theme_bw()

Lyft.price <- ggplot(rides %>% filter(cab_type == "Lyft" & name != "Taxi") %>% group_by(name), aes(reorder(as.factor(name), price, FUN = median), price, fill = name)) +
  geom_boxplot(position=position_dodge(1)) +
  labs(title = "Lyft", x = " ") +
  theme_bw()

grid.arrange(Uber.price, Lyft.price, nrow = 2)

```


# Distance vs Price
```{r}
ggplot(rides, aes(distance)) +
  geom_histogram(col = "black", fill="lightblue", alpha=0.5, position="identity", alpha=0.5) +
  theme_bw()

ggplot(rides %>% filter(cab_type == "Uber" & name != "Taxi"), aes(distance, price)) +
  geom_point(alpha = 0.05, aes(col = name)) +
  labs(title = "Uber") +
  theme_bw()

ggplot(rides %>% filter(cab_type == "Lyft" & name != "Taxi"), aes(distance, price)) +
  geom_point(alpha = 0.05, aes(col = name)) +
  labs(title = "Lyft") +
  theme_bw() 
```


# Price on weekdays:
```{r}
week.uber <- ggplot(rides %>% filter(cab_type == "Uber") %>% group_by(weekdays), aes(weekdays, price, fill = weekdays))+
  geom_violin(position=position_dodge(1)) +
  labs(title = "Uber", x = " ") +
  theme_bw()

week.lyft <- ggplot(rides %>% filter(cab_type == "Lyft") %>% group_by(weekdays), aes(weekdays, price, fill = weekdays))+
  geom_violin(position=position_dodge(1)) +
  labs(title = "Lyft", x = " ") +
  theme_bw()

grid.arrange(week.uber, week.lyft)
```


# Price on destination:
```{r}
dest.uber <- ggplot(rides %>% filter(cab_type == "Uber") %>% group_by(destination), aes(destination, price, fill = destination))+
  geom_boxplot(position=position_dodge(1)) +
  labs(title = "Uber", x = " ") +
  theme_bw()

dest.lyft <- ggplot(rides %>% filter(cab_type == "Lyft") %>% group_by(destination), aes(destination, price, fill = destination))+
  geom_boxplot(position=position_dodge(1)) +
  labs(title = "Lyft", x = " ") +
  theme_bw()

grid.arrange(dest.uber, dest.lyft)
```



```{r}


```

























